---
title: "Introduction to triageR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to triageR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(triageR)
```

## Overview

This vignette walks through a full clinical prediction modelling workflow using triageR, from raw data to a TRIPOD+AI-aligned report. We will use the **PIMA Indians Diabetes** dataset, a well-known real-world clinical dataset, to demonstrate each stage of the pipeline.

## 1. Load and prepare data

The PIMA dataset has a known data quality issue: several clinical measurements use `0` to represent a missing value, which we know is physiologically impossible for variables like blood pressure or BMI. We convert these to proper `NA`s before proceeding, this is a good example of the kind of cleaning `tr_load_clinical()` expects to receive data after.

```{r}
if (!requireNamespace("MASS", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
}
library(MASS)
pima <- Pima.tr2
pima$id <- seq_len(nrow(pima))

pima_loaded <- tr_load_clinical(pima, id_col = "id")
pima_loaded
```

## 2. Check and handle missing data

```{r}
tr_check_missing(pima_loaded)
```

`insulin` and `triceps` show substantial missingness — a realistic scenario in clinical datasets. We impute using `mice`:

```{r, message = FALSE, results = "hide"}
pima_for_impute <- pima_loaded[, setdiff(names(pima_loaded), "id")]
pima_imputed <- tr_impute(pima_for_impute, method = "mice", m = 5)
```

```{r}
sum(is.na(pima_imputed))
```

## 3. Fit a model

We split the data into training and test sets, then fit a logistic regression model.

```{r}
set.seed(42)
n <- nrow(pima_imputed)
train_idx <- sample(seq_len(n), size = floor(0.7 * n))

pima_train <- pima_imputed[train_idx, ]
pima_test  <- pima_imputed[-train_idx, ]

pima_model <- tr_fit(pima_train, outcome = "type", engine = "logistic_reg")
```

triageR also supports `"random_forest"` and `"boost_tree"` engines via the same interface — useful for comparing approaches without rewriting your pipeline:

```{r}
pima_model_rf  <- tr_fit(pima_train, outcome = "type", engine = "random_forest")
pima_model_xgb <- tr_fit(pima_train, outcome = "type", engine = "boost_tree")
```

## 4. Validate the model

Validation on a genuine holdout set produces discrimination metrics, a confusion matrix, and an ROC curve. We validate all three engines on the same holdout set to compare performance.

```{r}
pima_validation     <- tr_validate(pima_model, newdata = pima_test)
pima_validation_rf  <- tr_validate(pima_model_rf, newdata = pima_test)
pima_validation_xgb <- tr_validate(pima_model_xgb, newdata = pima_test)
```

```{r}
pima_validation$roc_plot
```

### Comparing engines

```{r}
engine_comparison <- dplyr::bind_rows(
  dplyr::mutate(pima_validation$metrics, engine = "logistic_reg"),
  dplyr::mutate(pima_validation_rf$metrics, engine = "random_forest"),
  dplyr::mutate(pima_validation_xgb$metrics, engine = "boost_tree")
)

tidyr::pivot_wider(engine_comparison, names_from = .metric, values_from = .estimate)
```

```{r}
pima_validation_rf$roc_plot
pima_validation_xgb$roc_plot
```

For the remainder of this vignette, we continue with the logistic regression model for explainability and reporting, as it offers the most directly interpretable coefficients for clinical use.

## 5. Explain the model

```{r}
tr_explain(pima_model, method = "permutation")
```

Glucose consistently emerges as the strongest predictor of diabetes in this dataset, consistent with established clinical literature.

## 6. Automated pipeline review

`tr_agent_review()` checks for common clinical modelling pitfalls: class imbalance, low events-per-variable ratio, possible data leakage, and small sample size.

```{r}
pima_review <- tr_agent_review(pima_train, pima_model, use_agent = FALSE)
```

## 7. Sensitivity analysis

```{r}
pima_sensitivity <- tr_sensitivity(pima_train, pima_model)
```

## 8. Generate a TRIPOD+AI-aligned report

Finally, all of the above can be compiled into a single reproducible report, aligned with TRIPOD+AI reporting guidance.

```{r, eval = FALSE}
tr_tripod_report(
  model = pima_model,
  validation = pima_validation,
  review = pima_review,
  sensitivity = pima_sensitivity,
  output_file = file.path(tempdir(), "pima_diabetes_report"),
  format = "html"
)
```

## Optional: AI-assisted method recommendation

If a Gemini API key is configured (see `?tr_recommend_method`), triageR can also suggest an appropriate statistical or ML approach based on the structure of your data:

```{r, eval = FALSE}
tr_recommend_method(
  pima_train,
  outcome = "type",
  context = "predicting diabetes onset in adult women"
)
```

## Summary

This vignette covered the full triageR workflow: loading and cleaning clinical data, handling missingness, fitting and validating a model, explainability, automated pipeline review, sensitivity analysis, and reproducible reporting, all using a real clinical dataset.
