---
title: "Clinical Prediction Model Report (TRIPOD+AI)"
format:
  html:
    default: true
  docx:
    default: false
params:
  bundle_path: NULL
---

```{r}
#| echo: false
#| message: false
bundle <- readRDS(params$bundle_path)
model <- bundle$model
model_type <- if (!is.null(bundle$model_type)) bundle$model_type else "classification"
validation <- bundle$validation
review <- bundle$review
sensitivity <- bundle$sensitivity
recommendation <- bundle$recommendation
is_survival <- model_type == "survival"
```

## Model Summary

```{r}
#| echo: false
#| results: asis
if (is_survival) {
  cat("- **Engine:**", model$engine, "\n")
  cat("- **Time column:**", model$time_col, "\n")
  cat("- **Event column:**", model$event_col, "\n")
  cat("- **Predictors:**", paste(model$predictors, collapse = ", "), "\n")
  cat("- **Training sample size:**", nrow(model$training_data), "\n")
  n_events <- sum(model$training_data[[model$event_col]] == 1, na.rm = TRUE)
  cat("- **Events observed:**", n_events, "\n")
} else {
  cat("- **Engine:**", model$engine, "\n")
  cat("- **Outcome:**", model$outcome, "\n")
  cat("- **Predictors:**", paste(model$predictors, collapse = ", "), "\n")
  cat("- **Training sample size:**", nrow(model$training_data), "\n")
}
```

## Method Recommendation (AI-assisted)

```{r}
#| echo: false
if (!is.null(recommendation)) {
  cat(recommendation)
} else {
  cat("Not run.")
}
```

## Validation Performance

```{r validation-performance}
#| echo: false
if (is_survival) {
  if (!is.null(validation)) {
    cat("**C-index:**", round(validation$c_index, 3),
        paste0("(SE: ", round(validation$c_index_se, 3), ")\n\n"))
    cat("**N:**", validation$n, " | **Events:**", validation$n_events, "\n\n")
    cat("C-index of 0.5 indicates no better than chance discrimination; ",
        "1.0 indicates perfect discrimination between patients who ",
        "experience the event sooner vs. later.")
  } else {
    cat("Not run.")
  }
} else {
  if (!is.null(validation)) {
    knitr::kable(validation$metrics)
  } else {
    cat("Not run.")
  }
}
```

```{r confusion-matrix}
#| echo: false
#| results: asis
if (!is_survival) {
  cat("### Confusion Matrix\n\n")
}
if (!is_survival) {
  if (!is.null(validation)) {
    print(knitr::kable(as.data.frame(validation$confusion_matrix$table)))
  } else {
    cat("Not run.")
  }
}
```

```{r calibration}
#| echo: false
#| results: asis
if (!is_survival) {
  cat("### Calibration\n\n")
}
if (!is_survival) {
  if (!is.null(validation) && !is.null(validation$calibration_plot)) {
    print(validation$calibration_plot)
  } else {
    cat("Not run.")
  }
}
```

*Validated on: `r if (!is.null(validation)) validation$validated_on else "N/A"`*

## Sensitivity Analysis

```{r}
#| echo: false
if (is_survival) {
  cat("Automated sensitivity analysis is not currently supported for survival models.")
} else if (!is.null(sensitivity)) {
  knitr::kable(tidyr::pivot_wider(sensitivity$comparison, names_from = .metric, values_from = .estimate))
} else {
  cat("Not run.")
}
```

## Pipeline Review Flags

```{r}
#| echo: false
if (!is.null(review) && nrow(review$flags) > 0) {
  knitr::kable(review$flags)
} else {
  cat("No issues flagged.")
}
```

```{r}
#| echo: false
if (!is.null(review) && !is.null(review$ai_summary)) {
  cat(review$ai_summary)
}
```

## TRIPOD+AI Compliance Notes

This report was generated using the `triageR` package. Please review the
following before submission:

- Confirm validation was performed on an independent dataset (see
  "Validated on" above — `"training_data"` indicates internal-only validation,
  which is NOT sufficient for TRIPOD+AI compliance).
- Confirm sample size and events-per-variable are adequate (see Pipeline
  Review Flags above).
`r if (!is_survival) "- Report calibration plots and confidence intervals alongside these point estimates for full TRIPOD+AI compliance."`
- This report is a drafting aid, not a substitute for the full TRIPOD+AI
  checklist, available at the EQUATOR Network.

*Report generated by triageR on `r Sys.time()`.*
